Epistemic Semantics in Guarded String Models

### 2020-06-17 tex/draft.tex
add formal detail to section 2

\caption{Epik program describing a possible-worlds event sequence model for two agents
  with information about one coin, and events of the agents semi-privately looking
at the coin.}
\label{epikfigure}
\end{figure}

@@

Rather than guarded strings, elements of a guarded string model for KAT are sets of
guarded strings.  In the application, these elements have the interpretation of
propositions (sets of possible worlds) and/or event types.  An event such as $u$ in
the guarded string model corresponds to the set of guarded strings where the bare
event is flanked by compatible truth assigments, as defined in an Epik program by the
event declaration and the constraint declaration. 
These have the status of event types, in that they
can ``happen'' in different possible worlds. Here happening corresponds to
incrementing the world with some compatible element of the event type.
Where $e$ is a bare event symbol, $\hat{e}$ is the
corresponding set of guarded strings,
consisting of the bare event decorated with compatible truth assignments.
See the examples in \ex{events1}.
% u (!q!p):(!qp) + (!qp):(q!p) + (q!p):(qp) 173
% d (!qp):(!q!p) + (q!p):(!qp) + (qp):(q!p)

\begin{example} \label{events1}
\(
\begin{array}[t]{l|c}
  e & \hat{e} \\ \hline  
  u & \{ 00u01, 01u10, 10u11\}\\
  d & \{ 01d00, 10d01, 11d10\}\\
  a_1 & \{10 a_1 10 \}\\
  a_0 & \{01 a_0 01\}\\
  b_1 & \{10 b_1 10 \}\\
  b_0 & \{01 b_0 01 \}\\
\end{array}
\)
\end{example}

Kleene Algebra with Tests has the algebraic signature
$\tuple{K,+,\cdot,*,\bar{},0,1}$.  To this we add an unary operation $\Diamond_a$ for
each agent, and a unary complent operation $\cdot^{c}$ on propositions.  This results
in the signature $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$ for
epistemic KAT with two agents.  In a guarded string model in our construction, the
elements of $K$ are sets of guarded strings.  The operation $+$ is set union. The
constant $0$ is the empty set.  The constant 1 is the set of all truth assignments
obeying the constraint on truth assignments, i.e. $\{10,10\}$ in the coin example.
The operation $\cdot$ is fusion product raised to sets: $x \cdot y$ is the set of all
defined fusion products of an element of $x$ with and element of $y$.  The operation
$*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$, where $x^i$ is the $i$-times
product of $x$ with itself and $x^0=1$.  Subsets of $1$ are also elements of $K$, and
these form the Boolean algebra of tests as a set Boolean algebra.  The overbar
operation is complement in this Boolean algebra.  The complement operation ${}^c$ is
complement at the level of sets of guarded strings, with $x^c = 0^c - x$, where the
operation on the right is set difference.  The epistemic modality $\Diamond_a$ is
interpreted using Kripke semantics, as pre-image relative to a fixed relation $R_a$
between guarded strings, $\Diamond_a x = \setabs{u}{\exists v.\element{v}{x} \wedge u
  R_a v}$.  Here $u$ and $v$ are guarded strings, while $x$ is an element of $K$.

@@

It remains to define the Kripke relation on guarded strings from an agent specification
as in Figure \ref{epikfigure}. An an agent specification pairs each bare event with a set of bare events,
and so determines a relation between bare events, call it relation $R_a$ for an
agent $e$.  This determines
a relation $\hat{R_a}$ between decorated events, see \ex{Rhat}.
It is generalized to a relation $\doublehat{R}$ on arbitrary guarded strings by
relational Kleene star, as defined in \ex{Rhathat}. The operation $\cdot$ in the
definition is KAT product, which enforces matching of tests.
The number of terms $n$ is construed as satisfying $n\geq 1$.
{\em What is desired for $n=0$?}

@@

@@

\section{Interpretation using lazy lists of strings}
%\pagebreak
%Hi Mom!

%\pagebreak

%Hi Sis!
%\pagebreak

@@

\begin{tabular}{lll}
propositions & $p,q...$ & \(p{:}{:}{=} \,
\begin{array}[t]{l}\tuple{\mbox{atomic events}}\,| \\
T \,|\, 
p{+}q \,|\,
p\mbox{\&}q \,|\, \\
p{;}q \;|\;
p{*} \;|\; {\sim}p \;|\\
{\diamond} a p
\end{array}\)
\end{tabular}

\caption{The language of Epik terms. Propositional terms are interpreted
  as sets of guarded strings.}
\label{termfigure}
\end{figure}

>>>>

\caption{Epik program describing a possible-worlds event sequence model for two agents
  with information about one coin, and events of the agents semi-privately looking
  at the coin.}
\label{fig:epik-example}
\label{epikfigure}
\end{figure}

@@

Rather than guarded strings, elements of a guarded string model for
KAT are sets of guarded strings.  In the application, these elements
have the interpretation of propositions (sets of possible worlds)
and/or event types.  An event such as $u$ in the guarded string model
corresponds to the set of guarded strings where the bare event is
flanked by compatible truth assigments, as defined in an Epik program
by the event declaration and the constraint declaration.  These have
the status of event types, in that they can ``happen'' in different
possible worlds. Here happening corresponds to incrementing the world
with some compatible element of the event type.  Where $e$ is a bare
event symbol, $\llbracket e \rrbracket^{\phi,P}$ is the corresponding
set of guarded strings, consisting of the bare event decorated with
compatible truth assignments denoted by $P$ that are admitted by the
world-constraints $\phi$.  See this formally in
Figure~\ref{fig:interp}, and the examples in \ex{events1}.
% u (!q!p):(!qp) + (!qp):(q!p) + (q!p):(qp) 173
% d (!qp):(!q!p) + (q!p):(!qp) + (qp):(q!p)

\begin{example} \label{events1}
\(
\begin{array}[t]{l|c}
  e & \llbracket e \rrbracket^{\phi, P} \\ \hline  
  u & \{ 00u01, 01u10, 10u11\}\\
  d & \{ 01d00, 10d01, 11d10\}\\
  a_1 & \{10 a_1 10 \}\\
  a_0 & \{01 a_0 01\}\\
  b_1 & \{10 b_1 10 \}\\
  b_0 & \{01 b_0 01 \}\\
\end{array}
\)
\end{example}

Kleene Algebra with Tests has the algebraic signature
$\tuple{K,+,\cdot,*,\bar{},0,1}$.  To this we add an unary operation
$\Diamond_a$ for each agent, and a unary complent operation
$\cdot^{c}$ on propositions.  This results in the signature
$\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$ for
epistemic KAT with two agents.

We interpret each term in as a set of guarded strings as indicated in
Figure~\ref{fig:interp}. A bare event $e$ is interpreted as the set of
guarded strings allowed by $\phi$ and $P$. The operation $p + q$ is
set union. The constant $0$ is the empty set.  The constant 1 is the
set of all truth assignments obeying the constraint on truth
assignments, i.e. $\{10,01\}$ in the coin example.  The operation
$\cdot$ is fusion product raised to sets: $x \cdot y$ is the set of
all defined fusion products of an element of $x$ with and element of
$y$.  The operation $*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$,
where $x^i$ is the $i$-times product of $x$ with itself and $x^0=1$.
Subsets of $1$ indicated by $t$ are also elements of $K$, and these
form the Boolean algebra of tests, denoting all atoms admitted by the
constraint $\phi$ and the boolean test $t$.  We ``overload'' $\cdot$
to further indicate boolean conjunction, $+$ to further indicate
boolean disjunction, and $0,1$ to indicate the zero and one elements
of the boolean algebra. The overbar operation denotes complement in
this Boolean algebra: $\overline t$ denotes all atoms $\alpha$ such
that $\alpha \Rightarrow \neg t$.  Other boolean operators (such as
implication $\Rightarrow$) are encoded as syntactic sugar in the
standard way. The complement operation ${p}^c$ is complement at the
level of sets of guarded strings, defined to every set the universe
$\_^*$ except those denoted by $p$. Finally, the epistemic modality
$\Diamond_a$ is interpreted using Kripke semantics, as pre-image
relative to a fixed relation $R_a$ between guarded strings,
$\Diamond_a x = \setabs{u}{\exists v.\element{v}{x} \wedge u R_a v}$.
Here $u$ and $v$ are guarded strings, while $x$ is an element of $K$.

@@

It remains to define the Kripke relation on guarded strings from an
agent specification as in Figure \ref{epikfigure}. An an agent
specification pairs each bare event with a set of bare events, and so
determines a relation between bare events, call it relation $R_a$ for
an agent $a$.  This determines a relation $\hat{R_a}$ between
decorated events, see \ex{Rhat}.  It is generalized to a relation
$\doublehat{R}$ on arbitrary guarded strings by relational Kleene
star, as defined in \ex{Rhathat}. The operation $\cdot$ in the
definition is KAT product, which enforces matching of tests.  The
number of terms $n$ is construed as satisfying $n\geq 1$.  {\em What
  is desired for $n=0$?}

@@

\begin{figure}
  \[\begin{array}{l >{\triangleq}c l}
  \llbracket 0 \rrbracket^{\phi,P} && \{\} \\  
  \llbracket 1 \rrbracket^{\phi,P} && \{\alpha \mid \alpha \Rightarrow \phi\}\\
  \llbracket e \rrbracket^{\phi,P}&& \{\alpha \cdot e \cdot \beta \mid
  (\alpha,\beta) \in P(e), \alpha, \beta \Rightarrow \phi \}
  \\ \llbracket \_ \rrbracket^{\phi,P}&& \bigcup_{e \in
    \dom(P)}\llbracket e \rrbracket^{\phi,P}\\
  \llbracket t \rrbracket^{\phi,P} &&  \{\alpha \mid \alpha \Rightarrow \phi \wedge t\}\\
  \llbracket{p+q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cup \llbracket q \rrbracket^{\phi,P}\\
  \llbracket{p \& q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cap \llbracket q \rrbracket^{\phi,P}\\
  \llbracket{p;q}\rrbracket^{\phi,P} && \llbracket p \rrbracket^{\phi,P} \cdot \llbracket{q}\rrbracket^{\phi,P}\\
  \llbracket p^*\rrbracket^{\phi,P} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\phi,P}\cdot \ldots \cdot \llbracket p \rrbracket^{\phi,P}}_{n}} \\
  \llbracket \sim p \rrbracket^{\phi,P} && \llbracket \_^*\rrbracket^{\phi,P} \setminus \llbracket p \rrbracket^{\phi,P} \\
  \llbracket \langle a \rangle~p \rrbracket && \doublehat{R_a}\llbracket p \rrbracket

  \end{array}
  \]
  \caption{Interpretation of Epik terms as sets of guarded strings}
  \label{fig:interp}
\end{figure}

@@

\section{Direct Bounded Implementation}

@@

\begin{tabular}{lll}
propositions & $p,q...$ & \(p{:}{:}{=} \,
\begin{array}[t]{l}\tuple{\mbox{atomic events}}\,|\,\_\, |\\
T \,|\, 
p{+}q \,|\,
p\mbox{\&}q \,|\, \\
p{;}q \;|\;
p{*} \;|\; {\sim}p \;|\\
{\langle a \rangle}~p
\end{array}\)
\end{tabular}

\caption{The language of Epik terms. Propositional terms are interpreted
  as sets of guarded strings.}
\label{fig:epik-grammar}
\label{termfigure}
\end{figure}
